
### Summary statistics

```{r Rm DEG, warning=FALSE, message=FALSE,  fig.height=8, fig.width=13}
# first create a table mapping the which define condition for which DE analysis has to be performed
  # recover the metadata from the "RawData"object
Meta=RawData$MetaData
  # remove replicate infos define the condition to look at and the corresponding reference Value
Meta2=Meta[!duplicated(Meta$Conditions),]
Cond=Meta2[,c("Conditions", DE_Contrast_Cond)]

  # Build the comparaison map using the "BuildDEComparaison" fonction
Map=BuildDEComparaison(Cond,"Conditions", Ref)
PMap=Map
PMap$Name=str_replace(PMap$Name, "_-_", " V ")

# Perform Differencial expression analysis
DETest=Genes_Diff_Expressed(NormData$Norm_Data,NormData$Norm_Data$targets$group_Cond,Group4InteractivePlot = data.frame(NormData$Norm_Data$targets[SecondaryComparaison]), AdditionalFactor=SecondaryComparaison, RefCond = Map$Ref[Map$Test_Cond==MainComparaison],CompareCond = Map$Test[Map$Test_Cond==MainComparaison],Method = "Paired",lfc_cuttoff = LFC_Cutoff, pvalue_cuttoff = pVal_Cutoff, Plots = F)


# Summary statistics
SumupBay=data.table::as.data.table(DETest$eBayes_Summary)
test=as.data.frame(dcast(SumupBay, V1~V2))
colnames(test)=str_replace(colnames(test), "-", " V ")
test%>%kable("html",caption = paste0("p-value<",as.character(pVal_Cutoff),", log2(FC)>",LFC_Cutoff)) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), font_size = 12, full_width = F)
```

Differential gene expression is performed by calculating the expression of in the `r unique(Meta[!Meta$Conditions%in%PMap$Ref[PMap$Test_Cond==MainComparaison],MainComparaison])` condition relative to the `r unique(Meta[Meta$Conditions%in%PMap$Ref[PMap$Test_Cond==MainComparaison],MainComparaison])` condition. The result is a log2(fold change) matrix indicating the expression of each genes relative to the reference condition. The various step described below use functions from the [limma](http://web.mit.edu/~r/current/arch/i386_linux26/lib/R/library/limma/html/01Introduction.html) package.

The first step consist in build a linear model to average read counts from each biological replicates. This operation is performed using the [lmFit()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/lmFit) function. The [makeContrasts()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/makeContrasts) function is then used to build a comparison matrix. Values produce by [lmFit()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/lmFit) and [makeContrasts()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/makeContrasts) are combined in another object using the [contrasts.fit()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/contrasts.fit). This procedure lead to the propagation of the statistics from the linear model. Finally the [ebayes()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/ebayes) function is used to extract from combined the object the relevant statistics of differential expression by empirical Bayes moderation of the standard errors towards a common value. 

To identify genes differentially expressed, the [decideTests()](https://www.rdocumentation.org/packages/limma/versions/3.28.14/topics/decideTests) and the [summary()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/summary) functions have been used. The first, classify the series of related t-statistics as up, down or not significant, while the second summaries the result of test. Differentially expressed genes has been identified using`r pVal_Cutoff` and `r LFC_Cutoff` as cutoff for pvalue and log2(FC) respectively. The tables presented in this slide show the summary of up and down regulated genes using these parameters.

***

### Relative expression of gene and clustering

<div class='left' style='float:left;text-align:center;width:50%;'>
```{r gene cluster , warning=FALSE, message=FALSE, fig.height=4, fig.width=8}

# Function to build violin plots
Generate_list_Stat_Compare=function(x){
  Cond=as.character(unique(DTfuseMetaNLFC[[x]]))

  ListComp=list()
  Ind=1
  for (i in 2:length(Cond)){
    ListComp[[Ind]]=as.character(c(Cond[1], Cond[i]))
    Ind=Ind+1
  }
  D=data.table::copy(DTfuseMetaNLFC[bol,])
  setnames(D, old = x, new = "Var")

  Formula=as.formula(paste("~", paste(Condlabel[Condlabel!=x], collapse = "+")))
  P=ggplot(D, aes(x=Var,y=Log.FC,fill=Var)) +
  geom_violin() +
  stat_compare_means( comparisons =ListComp , method = "t.test", label = "p.signif") +
  mynamesthemeSmall+ xlab(x)+ylab("Relative expr. [Log2(FC)]")+facet_wrap(Formula)+
   theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
  

  return(P)
}

Decide=decideTests(DETest$eBayes_Stat,adjust.method = "BH",p.value = pVal_Cutoff, lfc = LFC_Cutoff)

Bol=!(rowSums(abs(Decide)))<ncol(Decide)/Factor_Numb_Conditions_gene_diff_expressed
LFC=as.data.frame(DETest$eBayes_Stat$coefficients[Bol,])
pVal=as.data.frame(DETest$eBayes_Stat$p.value[Bol,])
# head(LFC)y§
PCA=M3C::umap(t(LFC))
PCA$layout=PCA$data

# ClustNumb=6
Dist=dist(PCA$layout,method = DistMethod)

Clust=cutree(hclust(Dist, method = AggloMethod),k=ClustNumb)

PCA_Data=as.data.frame(cbind(PCA$layout, Clust))
PCA_Data$Clust=as.factor(PCA_Data$Clust)

# merge PCA data with Gene info
PCA=merge(Genes, PCA_Data, by=0)
PCA$Note=str_wrap(PCA$Note,width=20)
# Plot The PCA scatter
P=ggplot(PCA, aes(x=X1, y = X2, color=Clust)) +
      geom_point(aes(Gene=Gene,orf19=ORF19_ID,Note=Note, Cluster=Clust))+
      ggtitle(paste(DistMethod,"distance \nwith" , AggloMethod, "agglomeration")) + 
  mynamesthemeSmall +
  scale_color_brewer(palette="Paired")+
  xlab("PC1") + ylab("PC2")
P=ggplotly(P)

## Build violin plots
# add cluster information to logFod change table
LFCnClust=cbind(LFC, Clust)

DTMeta=as.data.table(DETest$MetaData, keep.rownames = T)
DTLFC=as.data.table(LFCnClust,keep.rownames = T)
# reshape the LFC table
DTLFC=melt(DTLFC,id.vars = c("rn", "Clust"))

# fuse table adding condition and time to the LFC matrix
DTfuseMetaNLFC=DTMeta[DTLFC,on =.(rn=variable)]
# rename columns
setnames(DTfuseMetaNLFC, c("Comparaison",colnames(DETest$MetaData), "Gene", "Cluster","Log.FC"))

# Make the violin plot
Condlabel=colnames(DETest$MetaData)
bol=rep(T,nrow(DTfuseMetaNLFC))
Violin_plots=lapply(Condlabel, Generate_list_Stat_Compare)
# Violin_plots[[length(Violin_plots)+1]]=P

# Build the summary table
# add moderate t statisitc pvalue to the LFC table
LFCClustNpVal=cbind(LFCnClust,DETest$eBayes_Stat$F.p.value[Bol])
colnames(LFCClustNpVal)[ncol(LFCClustNpVal)]="Mod_Fstat_pVal"

# export the cluster table
# Add gene information to the LFC table
FullTble=merge(Genes,round(LFCClustNpVal,2),by="row.names")

ggarrange(plotlist=Violin_plots,align = "h")
P
# plotly::subplot(Violin_plots, nrows = 2,shareX = F,shareY = T,titleX = T, titleY = T,margin = c(0.1,0.1,0.1,0.1))
```

</div>

<div class='right' style='float:right;text-align:left;width:30%;'>

```{r plot heatmap, fig.height=8, fig.width=5}
Index=sort(as.numeric(PCA_Data$Clust), decreasing = T, index.return=T)
HC_Cluster=as.data.frame(Clust[Index$ix])
colnames(HC_Cluster)="HC_Cluster"

Annot=droplevels(DETest$MetaData)

Colorlist=brewer.pal(12, "Paired")
AnnotCol=sapply(colnames(Annot), function(x){
  RandNum=sample.int(12, length(levels(Annot[[x]])))
  Color=Colorlist[RandNum]
  names(Color)=levels(Annot[[x]])
  Color
}, simplify = FALSE,USE.NAMES = TRUE)
AnnotCol[["HC_Cluster"]]=Colorlist[1:length(levels(PCA_Data$Clust))]
names(AnnotCol[["HC_Cluster"]])=levels(PCA_Data$Clust)
# P1= pheatmap(as.matrix(LFC[Index$ix,]),cluster_cols = T, cluster_rows = F, main = "Canberra distance \nwith average agglomeration", width = 20, height = 60, annotation_row =HC_Cluster ,annotation_names_row = F ,annotation_col = DETest$MetaData, fontsize_row =2)
P1=pheatmap(as.matrix(LFC[Index$ix,]),cluster_cols = T, cluster_rows = F, main = paste(DistMethod,"\nwith",AggloMethod,"agglomeration") , width = 20, height = 60, annotation_row =HC_Cluster ,annotation_names_row = F ,annotation_col = Annot, fontsize_row =2, annotation_colors = AnnotCol)

```
</div>

***

### Summary table

```{r print summary table}
DT::datatable(FullTble[, c(2,3,7:ncol(FullTble))], filter = "top",extensions = 'Buttons', options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
  headerCallback = DT::JS(
    "function(thead) {",
    "  $(thead).css('font-size', '0.7em');",
    "}"
  )
), rownames = F)%>%
  DT::formatStyle(columns = colnames(FullTble[,c(2,3,7:ncol(FullTble))]), fontSize = '75%')

```

### Gene relative expression exploration

```{r Data exploration, warning=F}
# calculate gene coordinate in the genome 
Chr=Genes[, c("Chr", "Start", "End")]
Chr$Gene=rownames(Genes)
Chr$Chr=str_extract(Chr$Chr, Pattern_X)
Chr$ChrIdx=as.factor(Chr$Chr)
levels(Chr$ChrIdx)=1:length(levels(Chr$ChrIdx))
DT_Chr=as.data.table(Chr)
DT_Chr[, Max_Chr_Pos:=max(Start), by=ChrIdx]
DT_Chr_length=DT_Chr[,.(unique(Max_Chr_Pos)), by=ChrIdx][, Chr_Coord:=cumsum(V1)][,Chr_Start:=Chr_Coord-V1]
DT_Chr=DT_Chr[DT_Chr_length[,.(ChrIdx, Chr_Start)], on=.(ChrIdx=ChrIdx)][, Coord_Gene:=Start+Chr_Start]

# build a table containing all info for the interactivity 
FullTble=as.data.table(FullTble,keep.rownames = F)
Exploration1=DTfuseMetaNLFC[FullTble[, .(Row.names, Note)], on=.(Gene=Row.names)]
Exploration1=Exploration1[as.data.table(PCA_Data[,1:2],keep.rownames = T), on=.(Gene=rn)][DT_Chr[,.(Gene,Chr, Start, End,ChrIdx,Coord_Gene)], on=.(Gene=Gene)][!is.na(Comparaison),]
Exploration1[, MaxFC:= Log.FC[which.max(abs(Log.FC))], by=Gene]

# prepare shared data
SharedGroup=paste(c(NameSharedData,"Exp"), collapse = "_")
# SharedDT=SharedData$new(Exp, key = ~Gene)
SharedDT_PCA=SharedData$new(unique(Exploration1[, .(Gene, Note,X1,X2, Cluster, Log.FC)], by="Gene"), key = ~Gene, group=SharedGroup)
SharedDT_Violin=SharedData$new(Exploration1[, .SD,.SDcols= c("Gene","Cluster","Note",Condlabel, "Log.FC")], key = ~Gene, group=SharedGroup)
SharedDT_Filters=SharedData$new(unique(Exploration1[, .(Gene, Cluster, MaxFC), by="Gene"]), key = ~Gene, group=SharedGroup)

# Generate violin plots
Generate_list_Stat_Compare_list_plotly=function(x){
  Cond=as.character(unique(Exploration1[[x]]))

  ListComp=list()
  Ind=1
  for (i in 2:length(Cond)){
    ListComp[[Ind]]=as.character(c(Cond[1], Cond[i]))
    Ind=Ind+1
  }
  D=data.table::copy(Exploration1[bol,])
  setnames(D, old = x, new = "Var")

  Formula2=as.formula(paste("~", paste(Condlabel[Condlabel!=x], collapse = "+")))
  Formula1=as.formula(paste("~", paste(Condlabel[Condlabel==x], collapse = "+")))
  # P=ggplot(D, aes(x=Var,y=Log.FC,fill=Var)) +
  # geom_violin() +
  # stat_compare_means( comparisons =ListComp , method = "t.test", label = "p.signif") +
  # mynamesthemeSmall+ xlab(x)+ylab("Relative expr. [Log2(FC)]")+facet_wrap(Formula)+
  #  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
 P=SharedDT_Violin %>% 
   plot_ly() %>%
  add_trace(type="violin",x = Formula1,
          y = ~Log.FC,
          color = Formula2,
          marker = list( size = 5, opacity=0.5),
          points = "all",
          jitter = 0.7,
          pointpos = 0,
              hoverinfo = "text",
              hovertext = ~paste("Gene :", Gene,
                                "<br>Descritpion :", str_wrap(Note,20),
                                "<br>Cluster :", Cluster,
                                "<br>log2(FC) :", Log.FC)
  ) %>% layout( violinmode="group", legend=list(x=0, y=0.9))
  return(P)
}

Violin_plots=lapply(Condlabel, Generate_list_Stat_Compare_list_plotly)

p1=plot_ly(data = SharedDT_PCA) %>%
  add_markers(x = ~X1,
          y = ~X2, color=~as.factor(Cluster),
              hoverinfo = "text",
              hovertext = ~paste("Gene :", Gene,
                                "<br>Descritpion :", str_wrap(Note,20),
                                "<br>Cluster :", Cluster,
                                "<br>log2(FC) :", Log.FC))

Filters=list(
    filter_checkbox(paste(c(SharedGroup, "Cluster"), collapse  = "_"), "Cluster", SharedDT_Filters, ~Cluster, inline = TRUE),
    filter_slider(paste(c(SharedGroup, "Log_FC"), collapse  = "_"), "Highest Absolute log2(FC)",SharedDT_Filters, ~MaxFC, width = "100%",animate = F, round = 1, step = 0.5),
    filter_select(paste(c(SharedGroup, "Gene"), collapse= "_"), "Gene", SharedDT_Filters,~Gene)
  )

bscols(list(Filters,p1),Violin_plots)

# bscols(Filters,p1)

```

one two 

***
adsbvpashdvbpasuvbda


### Expression relative to the chromozome location

```{r, ste}
# prepare shared data
SharedGroup=paste(c(NameSharedData,"Chr"), collapse = "_")
# SharedDT=SharedData$new(Exp, key = ~Gene)
SharedDT_PCA=SharedData$new(unique(Exploration1[, .(Gene, Note,X1,X2, Cluster, Log.FC)], by="Gene"), key = ~Gene, group=SharedGroup)
SharedDT_Violin=SharedData$new(Exploration1[, .SD,.SDcols= c("Gene", "Note","Chr","Start", "End", "ChrIdx", "Coord_Gene",Condlabel, "Log.FC")], key = ~Gene, group=SharedGroup)
SharedDT_Filters=SharedData$new(unique(Exploration1[, .(Gene, Cluster, MaxFC), by="Gene"]), key = ~Gene, group=SharedGroup)


Generate_list_Stat_Compare_list_plotly_chr=function(x){
  Cond=as.character(unique(Exploration1[[x]]))

  ListComp=list()
  Ind=1
  for (i in 2:length(Cond)){
    ListComp[[Ind]]=as.character(c(Cond[1], Cond[i]))
    Ind=Ind+1
  }
  D=data.table::copy(Exploration1[bol,])
  setnames(D, old = x, new = "Var")

  Formula2=as.formula(paste("~", paste(Condlabel[Condlabel!=x], collapse = "+")))
  Formula1=as.formula(paste("~", paste(Condlabel[Condlabel==x], collapse = "+")))
  # P=ggplot(D, aes(x=Var,y=Log.FC,fill=Var)) +
  # geom_violin() +
  # stat_compare_means( comparisons =ListComp , method = "t.test", label = "p.signif") +
  # mynamesthemeSmall+ xlab(x)+ylab("Relative expr. [Log2(FC)]")+facet_wrap(Formula)+
  #  theme(legend.position="none",axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
 P=Exploration1 %>% 
   plot_ly() %>% 
  group_by(ChrIdx)%>%
  summarise(MinX=min(Coord_Gene), MaxX=max(Coord_Gene), X=unique(Chr), MinLFC=min(Log.FC))%>%
  layout( legend = list(orientation = "v",   # show entries horizontally
                     xanchor = "center",  # use center of legend as anchor
                     x = 1, y=1),
          xaxis=list(
              showticklabels = FALSE,
              tickvals=~MinX,
              gridwidth=4,
              gridcolor=toRGB("blue", alpha = 0.6)) ,
          annotations= list(x=~(MaxX-(MaxX-MinX)/2),
          y=~min(MinLFC),
          text=~X,
          ay=20,
          ax=0))%>%
  add_markers(data=SharedDT_Violin,
              x = ~Coord_Gene,
              y = ~Log.FC, 
              color=Formula2,
              symbol = Formula1,
              hoverinfo = "text",
              hovertext = ~paste("Gene :", Gene,
                                "<br>Descritpion :", str_wrap(Note,20),
                                "<br>Chromosome :", Chr ,
                                "<br>Start :" , Start,
                                "<br>End :", End))
  return(P)
}

Violin_plots=lapply(Condlabel, Generate_list_Stat_Compare_list_plotly_chr)


p1=plot_ly(data = SharedDT_PCA) %>%
  add_markers(x = ~X1,
          y = ~X2, color=~as.factor(Cluster),
              hoverinfo = "text",
              hovertext = ~paste("Gene :", Gene,
                                "<br>Descritpion :", str_wrap(Note,20),
                                "<br>Cluster :", Cluster,
                                "<br>log2(FC) :", Log.FC))

Filters=list(
    filter_checkbox(paste(c(SharedGroup, "Cluster"), collapse  = "_"), "Cluster", SharedDT_Filters, ~Cluster, inline = TRUE),
    filter_slider(paste(c(SharedGroup, "Log_FC"), collapse  = "_"), "Highest Absolute log2(FC)",SharedDT_Filters, ~MaxFC, width = "100%",animate = F, round = 1, step = 0.5),
    filter_select(paste(c(SharedGroup, "Gene"), collapse= "_"), "Gene", SharedDT_Filters,~Gene)
  )

bscols(Filters,p1)
bscols(Violin_plots)

```
blalb